### 2021-08-21 plos_latex_template.tex
accepted most changes from Matt, Heather, Janani, Andrea

% % Please keep the Author Summary between 150 and 200 words
% % Use first person. PLOS ONE authors please skip this step. 
% % Author Summary not valid for PLOS ONE submissions.   
% \section*{Author summary (I don't think we need it)}
% Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

@@

@@

%some groups face discrimination and might be underrepresented in your community and event \cite{timperley_he_2020}

\section*{Rule 1: Set realistic but bold diversity indicators}

As a starting point, identify which groups are usually unsupported or discriminated against in the community or event of interest, and determine the contributing factors. Next, set measurable indicators of success when organizing a diverse conference. The numbers resulting from these indicators are not the primary goal, and this is not about performative allyship. Instead, these numbers should guide and help hold the organizing team accountable towards these diverse and inclusive goals.

@@

\section*{Rule 2: Have an inclusive and diverse organizing team}
A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team. Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity. Particular attention should be paid to the usually underrepresented groups (see \textbf{Rule 1} discussing the dimensions of diversity lacking in your community and event). To ensure a deeper understanding of the challenges in different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. Even if creating and maintaining such an interdisciplinary team is challenging, the positive outcomes far outweigh any minor inconveniences, and the team will grow, learn, and embrace more inclusive practices. For instance, if the organizing team is global and remote, time zones can be hard to manage, with meetings scheduled at odd hours. The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best platforms to mechanisms to account for every culture and situation. 
A successful, inclusive team will not only be a safe space for people to learn and grow but also may help identify critical aspects that are working or failing to improve inclusiveness and accessibility together \cite{hong_groups_2004}.

@@

While this rule may seem intuitive, one would be surprised by how often conference organizers tend to do everything themselves instead of reaching out to the right people for the right task -- likely because the latter involves careful planning and communication well ahead of time or because they naively overestimate their expertise in areas they lack practical experience in. For e.g., a cisgender or non-neurodiverse person estimating what might work best for the LGBTQ+ or people with hearing and seeing difficulty. 
Therefore, to build your team of organizers for the conference, you need to actively seek out and recruit the best people for each task, akin to filling your team with the best, diversified portfolio. However, the intention to put together a diverse team alone is not enough. One needs to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions [\textbf{Fig. 1}]. To achieve this, it will be necessary to go beyond one's limited networks when inviting people to the organizing team, keynote speakers, program committee, session chairs, reviewers, and others. Ensuring diversity in each of these teams and panels is a deliberate process -- if there is not much diversity in the first set of names, it needs to be revised to include people from minoritized groups. If these groups aren't showing up in the A-list, it is seldom because of their technical acumen -- it is because we, as organizers, are not used to turning every work and looking in unfamiliar places to bring in the talent.  Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyer_notice_2021, sarabipour_evaluating_2020}, thus encouraging us to look further to find great people that are not routinely in the spotlight. We need to go beyond our narrow and often limited networks. So, as long as we start with an inclusive and diverse organizing team (\textbf{Rule 2}), we can aim at finding the right people for various aspects of the conference organization. The regional and local communities are also good sources to tap into. For example, for useR! 2021, groups like AfricaR, R-Ladies, MiR, R Forwards, and LatinR were fundamental to reach people for the organizing team, potential attendees, and sponsors.

Disabled people often repeat "Nothing about us without us". This means the actual life experiences of people from URM are not replaceable with good intentions from people outside these groups. The same holds for other dimensions of diversity. No space will be truly inclusive and welcoming to URM without people from these groups -and their experiences-. This means people in privileged groups may need to step down from positions of power. Of course, no-one has the same experiences and experiences the same oppressions in life. When building a diverse team, bear in mind that

@@

Conferences, even virtual ones, should have registration fees for two primary reasons. The first one is that preparing the conference involves a lot of concerted effort and costs money (e.g., commercial registration tools, captioning, or conference venue, if in-person) and this translates to the participation fee. The second reason has to do with the human psychology -- people value more the things they pay for, and there is a lower attendance rate for free events with no registration costs \cite{eventbrite_ultimate_2017}. On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates. Usually, there is a higher fee for people from the industry than for academia (~ proportionate to income). We should also consider a lower fee for non-profit organizations, government employees, or freelancers, and use conversion factors (for country of residence using data from International Comparison Program report of the World Bank \cite{arend_disparity_2019}). It is important to include discounts for students as well as for postdocs (early career trainees) to encourage their participation \cite{sarabipour_evaluating_2020, andalib_postdoc_2018, kaplan_postdoc_2012}. Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, their payment category should be explicitly defined. Disabilities may also increase the overall cost of conferences, especially in-person \cite{de_picker_rethinking_2020, irish_increasing_2020}; discounts would help reduce the financial burden for these factions too.

Representation is an important aspect of diversity and inclusivity: see people like you, from your country, that speak your language is one of the best way to feel you belong, this is way the rates can not only being different for registration, but also, for sponsors, this allows us to have institutions and companies from all over the world being part of the conference which also means they will communicate about our event in their local communities.

@@

Conferences are usually inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}. Each of these groups should be considered at every stage of the planning process. Welcoming representatives from these groups into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule 2}). If the conference is in person, the venue must be accessible for people with mobility limitations. A quiet space for neurodivergent people and a space for breastfeeding should also be provided. Images used in communications about the conference should be accompanied by alternative text and videos with captions and sound. Platforms for conference registration, abstract submission -- and chat platforms, if the conference is virtual or hybrid -- should be screen-reader friendly and with low technology requirements (that do not have compatible hardware, software, and internet connection requirements to function, all this tested beforehand). Captioning for presentations should be available in more than one language if possible. If the conference is in-person, presenters should speak into a microphone to make it easier for captioners or interpreters to listen to them. Accessibility guidelines for slides and presentations should be provided to the presenters, and their use should be encouraged (see \url{https://user2021.r-project.org/participation/accessibility/} for example). Among other things, the speakers should provide raw and accessible material to their talks before the conference, e.g., such as RMarkdown, HTML, or TeX files; if they prerecord their presentation, they should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed. Accessibility awards at the end of the conference are an excellent way to acknowledge the speakers who were mindful of inclusiveness when preparing and delivering their talks.

@@

\section*{Rule 8: Have an important online component of the conference}

@@

In international conferences, English is often the official language. Submissions, presentations, tutorials, and workshops are in English. The platforms, the webpage, and official communications are also in English. While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{niner_better_2021}. Excluding them may potentially lead to missing innovative contributions brought about by a language barrier. Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier. For instance, hosting one international day/session per conference might be a great place to start!

@@

To ensure an inclusive, friendly, and safe space in the conference, the organizers need to a clearly define the code of conduct and a team to enforce it \cite{favaro_your_2016}. `The purpose of a code of conduct is to protect members of a community from harm in that community's spaces. The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.' \cite{aurora_how_2018}
The code of conduct should be displayed prominently in several spaces of the conference to deter people from incurring in unaccepted behavior.

The code of conduct team should receive training to understand unambiguously what to and not to enforce in a code of conduct, how to receive reports, respond to incidents, and communicate their responses. A diverse code of conduct team will be more understanding of intersectionality issues in discrimination and harassment practices.

@@

The ten rules stated here can be adapted depending on the conference format and settings. From our own experience, we are aware that enforcing these rules requires a tremendous amount of work and is not compensated financially or otherwise. However, the rewards from these suggested best practices will provide a healthier, stronger, and more inclusive community, empowering historically marginalized people and learning from them, thereby increasing the overall quality of the conference. Also, since these are early days and accessibility is not as widespread as it should be, the process might seem exacting. Hopefully, if more conferences and domains employ these techniques, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead. After all, the strength of any community lies in its ability to be adaptive, inclusive, and accessible.

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 [insert full list of organizers \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for their financial support for useR! 2021.

>>>>

% % Please keep the Author Summary between 150 and 200 words

@@

\section*{Rule 1: Embrace all dimensions of diversity}

The first step towards a diverse and inclusive conference is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event \cite{timperley_he_2020}.

@@

As a starting point, identify which groups are usually unsupported or discriminated against in the community or event of interest, and determine the contributing factors. Next, think about set measurable indicators of success when organizing a diverse conference. The numbers resulting from these indicators are not the primary goal. Instead, this vision should guide and help hold the organizing team accountable towards these diverse and inclusive goals.

@@

\section*{Rule 2: Have an inclusive and diverse organizing team}
A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team. Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity. Particular attention should be paid to the usually underrepresented groups (see \textbf{Rule 1} discussing the dimensions of diversity lacking in your community and event). To ensure a deeper understanding of the challenges in different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. Even if creating and maintaining such an interdisciplinary team is challenging, the positive outcomes far outweigh any minor inconveniences, and the team will grow, learn, and embrace more inclusive practices. For instance, if the organizing team is global and remote, time zones can be hard to manage, with meetings scheduled at odd hours. The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best ways to account for every culture and situation. 
An inclusive team will let everyone know if the organization is failing in any aspect of inclusiveness or accessibility, and help improve it \cite{hong_groups_2004}.

@@

While this rule may seem intuitive, one would be surprised by how often conference organizers tend to do everything themselves instead of reaching out to the right people for the right task -- likely because the latter involves careful planning and communication well ahead of time or because they naively overestimate their expertise in areas they lack practical experience in. For example, a cisgender or abled person estimating what might work best for transgender people or people with sensory disabilities. 
Therefore, to build your team of organizers for the conference, you need to actively seek out and recruit the best people for each task. The intention to put together a diverse team alone is not enough. One needs to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions [\textbf{Fig. 1}]. To achieve this, it will be necessary to go beyond one's limited networks when inviting people to the organizing team, keynote speakers, program committee, session chairs, reviewers, and other roles. Ensuring diversity in each of these teams and panels is a deliberate process -- if there is not much diversity in the first set of names, it needs to be revised to include people from minoritized groups. If these groups aren't showing up in the A-list, it is seldom because of their technical acumen -- it is because we, as organizers, are not used to turning every work and looking in unfamiliar places to bring in the talent.  Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyer_notice_2021, sarabipour_evaluating_2020}, thus encouraging us to look further to find great people that are not routinely in the spotlight. We need to go beyond our narrow and often limited networks. So, as long as we start with an inclusive and diverse organizing team (\textbf{Rule 2}), we can aim at finding the right people for various aspects of the conference organization. The regional and local communities are also good sources to tap into. For example, for useR! 2021, groups like AfricaR, R-Ladies, MiR, Forwards, and LatinR were fundamental to reach people for the organizing team, potential attendees, and sponsors.

Disabled people often repeat "Nothing about us without us". This means the actual life experiences of people from URM are not replaceable with good intentions from people outside these groups. The same holds for other dimensions of diversity. No space will be truly inclusive and welcoming to URM without people from these groups bringing their experience to bear. This means people in privileged groups may need to step down from positions of power. Of course, no-one has the same experiences and experiences the same oppressions in life. %When building a diverse team, bear in mind that

@@

Conferences, even virtual ones, should have registration fees for two primary reasons. The first one is that preparing the conference involves a lot of concerted effort and costs money (e.g., commercial registration tools, captioning, or conference venue, if in-person) and this translates to the participation fee. The second reason has to do with psychology -- people value more the things they pay for, and there is a lower attendance rate for free events with no registration costs \cite{eventbrite_ultimate_2017}. On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates. Usually, there is a higher fee for people from the industry than for academia ($\sim$ proportionate to income). We should also consider a lower fee for non-profit organizations, government employees, or freelancers, and use conversion factors (for country of residence using data from International Comparison Program report of the World Bank \cite{arend_disparity_2019}). It is important to include discounts for students as well as for postdocs (early career researchers) to encourage their participation \cite{sarabipour_evaluating_2020, andalib_postdoc_2018, kaplan_postdoc_2012}. Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, their payment category should be explicitly defined. Disabilities may also increase the overall cost of conferences, especially in-person \cite{de_picker_rethinking_2020, irish_increasing_2020}; discounts would help reduce the financial burden for these factions too.

Representation is an important aspect of diversity and inclusivity: seeing people like you, from your country, that speak your language is one of the best way to feel you belong, this is why the rates can not only be different for registration, but also, for sponsors, this allows us to have institutions and companies from all over the world being part of the conference which also means they will communicate about our event in their local communities.

@@

Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}. People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status. Each of these groups should be considered at every stage of the planning process. Welcoming representatives from these groups into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule 2}). If the conference is in person, the venue must be accessible for people with mobility limitations. Accommodations such as a quiet space for neurodivergent people and a space for breastfeeding should also be provided. Images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound. Platforms for conference registration and abstract submission, websites and chat platforms should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection). These aspects should be tested well in advance of going live. Captioning for presentations should be available in more than one language if possible. If the conference is in-person, presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them. Accessibility guidelines for slides and presentations should be provided to the presenters, and their use should be encouraged (see \url{https://user2021.r-project.org/participation/accessibility/} for example). Among other things, the speakers should provide raw and accessible material to their talks before the conference, e.g., R Markdown, HTML, or \TeX{} files; if they prerecord their presentation, they should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed. Adding captions to recorded videos is relatively easy these days and is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.  Accessibility awards at the end of the conference are an excellent way to acknowledge the speakers who were mindful of inclusiveness when preparing and delivering their talks.

@@

\section*{Rule 8: Have a strong online component of the conference}

@@

In international conferences, English is often the official language. Submissions, presentations, tutorials, and workshops are in English. The platforms, the webpage, and official communications are also in English. While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{niner_better_2021}. Excluding them may potentially lead to missing innovative contributions due to a language barrier. Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier. For instance, hosting one international day/session per conference might be a great place to start!

@@

To ensure an inclusive, friendly, and safe space in the conference, the organizers need to adopt a code of conduct and set up a team to enforce it \cite{favaro_your_2016}. The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora_how_2018}.%The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.' 
The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.

The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses. A diverse code of conduct team will be more understanding of intersectionality issues in discrimination and harassment practices.

@@

The ten rules stated here can be adapted depending on the conference format and settings. From our own experience, we are aware that enforcing these rules requires a tremendous amount of work that is generally not compensated financially or otherwise. However, the rewards from these suggested best practices will provide a healthier, stronger, and more inclusive community, empowering historically marginalized people and learning from them, thereby increasing the overall quality of the conference. Since these are early days and accessibility is not as widespread as it should be, the process might seem exacting. Hopefully, if more conferences and domains employ these techniques, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead. After all, the strength of any community lies in its ability to be adaptive, inclusive, and accessible.

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 [ \url{https://user2021.r-project.org/about/global-team/}] for their valuable contribution to an inclusive conference experience, and the R Foundation for charging us with the organization of useR! 2021 and supporting us through the process.
